# Activity Monitoring

Download the file and load the data into a dataframe:
```{r load, echo=TRUE}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip","repdata_data_activity.zip")
unzip("repdata_data_activity.zip")
activity <- read.csv("activity.csv", sep=",", header=TRUE)
```

### Histogram of Total Number of Steps taken per day:

```{r steps_per_day, echo=TRUE}
steps_per_day <- aggregate(activity$steps, by = list(activity$date), "sum", na.rm=TRUE)
colnames(steps_per_day) <- c("Date", "Total_Steps")
plot <- hist(steps_per_day$Total_Steps, xlab = "Number of Steps", ylab = "Frequency", main= "Total steps taken per day")

```

### Mean and Median of the total number of steps taken per day:


```{r Mean, echo=TRUE}
mean_steps <- mean(steps_per_day[,"Total_Steps"], na.rm=TRUE)
mean_steps
median_steps <- median(steps_per_day[,"Total_Steps"], na.rm = TRUE)
median_steps
```

### Daily Average Activity Pattern

```{r average, echo=TRUE}
average_steps <- aggregate(activity$steps, by = list(activity$interval), "mean", na.rm=TRUE)
colnames(average_steps) <- c("Interval", "Steps")
plot(average_steps$Interval, average_steps$Steps, xlab = "Interval", ylab="Average Number of Steps", main = "Average Number of Steps for each Time Interval", type="l")

```

### Interval with the maximum number of steps:
```{r max_interval, echo=TRUE}
max_steps <- subset(average_steps, average_steps$Steps==max(average_steps$Steps), select = Interval)
max_steps
```

### Input Missing Values

The number of missing values in the dataset are:
```{r missing, echo=TRUE}

## Generates a dataframe with all rows that have missing values
missing_values <- subset(activity, is.na(activity$steps))
nrow(missing_values)
```

New dataset with missing data filled with mean for the 5 minute interval:


```{r, echo=TRUE}

## Merge 'missing values' dataframe with 'average steps taken in a day' dataframe by the interval
complete <- merge(missing_values, average_steps, by.x = "interval", by.y="Interval")
complete_final <- subset(complete, select=c("Steps", "date", "interval"))
colnames(complete_final) <- c("steps", "date", "interval")

## Original dataframe without the missing values
activity_new <- subset(activity, !is.na(activity$steps))

## Combine both dataframes together
final <- rbind(activity_new, complete_final)

## Dataframe with totals summed
steps <- aggregate(final$steps, by=list(final$date), "sum")
colnames(steps) = c("date","steps")

```

### Histogram of total number of steps taken each day
```{r, echo=TRUE}
hist(steps$steps, xlab = "Number of Steps", ylab ="Frequency", main = "Total number of Steps taken each day")
mean_steps_final <- mean(steps[,"steps"])
mean_steps_final
median_steps_final <- median(steps[,"steps"])
median_steps_final

```

### Difference in activity patterns between weekdays and weekends:

```{r}
final$day <- weekdays(as.Date(final$date))
final$weekday_weekend <- factor(final$day, c("Weekday", "Weekend"))
final$weekday_weekend <- ifelse(final$day == "Sunday" | final$day == "Saturday", "Weekend", "Weekday")
```

Average Number of Steps for weekdays and weekends

```{r plot, echo=TRUE}
library(ggplot2)

average_steps_final <- aggregate(final$steps, by = list(final$interval, final$weekday_weekend), "mean")
colnames(average_steps_final) <- c("interval", "day", "steps")
ggplot(average_steps_final, aes(average_steps_final$interval, average_steps_final$steps)) + facet_grid(average_steps_final$day~.) + geom_line() + labs(title = "Average Number of Steps for each interval (Weekday/Weekend)", x = "Interval", y="Average Number of Steps")
```








